---
title: "Challenge 1 Simulation and Correlation"
author: "Xiang (Sean) Cheng (chengxiang.cn@gmail.com)"
date: "June 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Revenue growth prediction is an interesting and practical problem. I do have questions because I am new to this type of revenue prediction problem, but I will make some assumptions due to limitted time. In real working environment, we need to communciate to make decisions about the assumptions and parameters.
([More information about the problem](http://htmlpreview.github.com/?https://raw.githubusercontent.com/seanxiangcheng/datachallenge/master/Problemsets/Data_Challenge_Guide.html))

#### Assumptions
  1. The question we want to answer is the dependence of total revenue growth on correlation.
  2. Correlation can be positive or negative;
    * costomers need both features: positive correlation;
    * costomers need only 1 feature: negative or no correlation;
  3. Both features boost the revenue cumulatively (after correlation);
    * The product with more features has higher price;
    * Each feature can generate regular cash flow (by selling disposable consumables) ;
  4. Both features are unlikely to decrease revenue.
    * Negative growth rate has very small or no probability;
    * Normal distribution: standard deviation<=0.02;
    * Uniform distribution: min>=0.01;
  5. Correlations are in a reasonable range. Here we assume it is between -0.8 and 0.8.
  
## Task 1: same distributions
In this task, for simplicity, assume both are from the same distribution, either a normal distribution or a uniform distribution. 

#### 1. Normal distributions

```{r, echo=FALSE, warning=FALSE}
# load libraries and constants
library(MASS)
library(ggplot2)
r.min = -0.8
r.max = 0.8
```

The 2 functions to generate correlated random normal variables and to examine the mean and correlation matrix.
```{r, warning=FALSE}
# function to generate correlated random normal variables
normal_xy <- function(r=0.5, n=1e3, mu=c(0.05, 0.07), var=c(0.02^2, 0.02^2)){
  sigma = matrix(c(var[1], r*prod(sqrt(var)), r*prod(sqrt(var)), var[2]),nrow=2)
  normal_xy = mvrnorm(n=n, mu=mu, Sigma=sigma)
  normal_xy = data.frame(normal_xy)
  names(normal_xy) = c('X', 'Y')
  return(normal_xy)
}
# function to check the mean and correlation of the random variables
check_mean_cor <- function(df_xy){
  mus = apply(df_xy, 2, mean)
  cat("mean of X and Y:\n"); print(mus)
  sigma = cor(df_xy)
  cat("\ncorrelation matrix:\n"); print(sigma)
}
```

A quick check of the 2 functions:
```{r, warning=FALSE}
check_mean_cor(normal_xy(r=0.5, n=1e5))
```

Here, we simulate the revenue and plot the 90% confidence interval of the growth rate with both features. The standard deviations are 0.02 and 0.02, respectively.
```{r, warning=FALSE, message=FALSE}
# 100 different correlations
n.obs = 100; CI=0.9;
rs = seq(r.min, r.max, length.out=n.obs)
total.mean = rep(0, n.obs)
CI.low = rep(0, n.obs)
CI.high = rep(0, n.obs)
for(i in 1:n.obs){
  r = rs[i]
  total = rowSums(normal_xy(r=r, n=1e4)) 
  total.mean[i] = mean(total)
  CI.low[i] = quantile(total, (1-CI)/2.)
  CI.high[i] = quantile(total, (1+CI)/2.)
}
df = data.frame(rs=rs, mean=total.mean, low=CI.low, high=CI.high)
ggplot(df, aes(x=rs , y = mean)) + geom_point(size = 1) + geom_line()+
  geom_line(aes(y = low, colour = "red"))+  geom_line(aes(y = high, colour = "red"))+
  labs(x="Correlation between X and Y",y="Total Revenue Growth Rate", title=paste0("Total Growth Rate (Confidence Interval ", CI*100, "%)") )+ theme(legend.position="none")
```


#### 2. Uniform distributions
In a similar manner, we can also get the revenue prediction of uniformly distributed rates. Instead of using a build-in function, we need to use cholesky decomposition to transform the independent random variables to correlated ones. (Functions are hidden in raw R markdown file by setting 'echo=FALSE'.)
The ranges of uniform distribtuions are [0.03, 0.07] and [0.05, 0.09], respectively.

```{r, warning=FALSE, echo=FALSE}
# function to generate correlated random normal variables
uniform_xy <- function(r=0.5, n=1e3, mu=c(0.05, 0.07), span=c(0.02, 0.02)){
  unif_xy = rbind(runif(n, mu[1]-span[1], mu[1]+span[2]), runif(n, mu[2]-span[2], mu[2]+span[2]))
  #print(rowMeans(unif_xy))
  sigma = matrix(c(1, r, r, 1),nrow=2)
  U = t(chol(sigma))
  unif_xy = U %*% unif_xy
  unif_xy = data.frame(t(unif_xy))  
  names(unif_xy) = c('X', 'Y')
  return(unif_xy)
}

# 100 different correlations
n.obs = 100; CI=0.9;
rs = seq(r.min, r.max, length.out=n.obs)
total.mean = rep(0, n.obs)
CI.low = rep(0, n.obs)
CI.high = rep(0, n.obs)
for(i in 1:n.obs){
  r = rs[i]
  total = rowSums(uniform_xy(r=r, n=1e4)) 
  total.mean[i] = mean(total)
  CI.low[i] = quantile(total, (1-CI)/2.)
  CI.high[i] = quantile(total, (1+CI)/2.)
}

df = data.frame(rs=rs, mean=total.mean, low=CI.low, high=CI.high)
ggplot(df, aes(x=rs , y = mean)) + geom_point(size = 1) + geom_line()+
  geom_line(aes(y = low, colour = "red"))+  geom_line(aes(y = high, colour = "red"))+
  labs(x="Correlation between X and Y",y="Total Revenue Growth Rate", title=paste0("Total Growth Rate (Confidence Interval ", CI*100, "%)") )+ theme(legend.position="none")
```


## Task2: different distributions
We here assume the impact of X is normally distributed while the impact of Y is uniformally distributed. The functions are hidden in the raw R markdown file ('echo=FALSE').

```{r, warning=FALSE, echo=FALSE}
# function to generate correlated random normal variables
rand_xy <- function(r=0.5, n=1e3, mu=c(0.05, 0.07), varx=0.02^2, spany=0.02){
  rxy = rbind(rnorm(n, mu[1], sqrt(varx)), runif(n, mu[2]-spany, mu[2]+spany))
  #print(rowMeans(unif_xy))
  sigma = matrix(c(1, r, r, 1),nrow=2)
  U = t(chol(sigma))
  rxy = U %*% rxy
  rxy = data.frame(t(rxy))  
  names(rxy) = c('X', 'Y')
  return(rxy)
}
```

Test the correlation between the 2 correlated samples.
```{r}
check_mean_cor(normal_xy(r=0.5, n=1e5))
```


```{r, warning=FALSE, echo=FALSE}
# 100 different correlations
n.obs = 100; CI=0.9;
rs = seq(r.min, r.max, length.out=n.obs)
total.mean = rep(0, n.obs)
CI.low = rep(0, n.obs)
CI.high = rep(0, n.obs)
for(i in 1:n.obs){
  r = rs[i]
  total = rowSums(rand_xy(r=r, n=1e4)) 
  total.mean[i] = mean(total)
  CI.low[i] = quantile(total, (1-CI)/2.)
  CI.high[i] = quantile(total, (1+CI)/2.)
}
df = data.frame(rs=rs, mean=total.mean, low=CI.low, high=CI.high)
ggplot(df, aes(x=rs , y = mean)) + geom_point(size = 1) + geom_line()+
  geom_line(aes(y = low, colour = "red"))+  geom_line(aes(y = high, colour = "red"))+
  labs(x="Correlation between X and Y",y="Total Revenue Growth Rate", title=paste0("Total Growth Rate (Confidence Interval ", CI*100, "%)") )+ theme(legend.position="none")
```


__Note__:

  1. **Correlated random variables**:
  In my research, we use advanced Monte Carlo algorithms, complex networks, and C++ to simulate physical systems. Mostly, we use independent random variables, so I don't have much experience in generating correlated random variables using R. There may be better ways to do, and I am very intestered to learn. I do hope I can learn new things from a job. The good thing is that I have a soild foundation in statistics, math, and programming. It should be quick for me to learn new things.

  2. **Programming languages**:
  I can use R, Python, and C++. R is chosen here to be consistent with your choice of generating the guide files. I also have experience in interpreting outputs from SAS, STATA, and SPSS. My diverse skill set may make team-work easier.